---
title: "KV6015-DataAnalyticsRAssessmentAnswer"
author: "Rayhan Chowdhury Bijoy"
format: 
  html:
      self-contained: true
date: "2024-12-05"
editor: source
---

```{r}
library(readxl)
library(tibble)
library(dplyr)
library(ggplot2)
library(scales)
library(tidyr)
```

# Load the data
```{r}
# Load the data from the Excel file
data <- read_excel("./PreparedData.xlsx")

# Convert the data to a tibble
data_tibble <- as_tibble(data)

# View the tibble (first few rows)
print(data_tibble)
```

#-----------------------1.	Hypothesis 01 (Gender-based Focus): Linear Model Test-----------------------
```{r}
# Create a 'Period' column based on the 'Date'
data_tibble <- data_tibble %>%
  mutate(Period = case_when(
    Date %in% c("2018/19", "2019/20") ~ "Pre-pandemic",
    Date %in% c("2020/21", "2021/22") ~ "Pandemic",
    Date %in% c("2022/23", "2023/24") ~ "Post-pandemic",
    TRUE ~ "Unknown"
  )) %>%
  mutate(Period = factor(Period, levels = c("Pre-pandemic", "Pandemic", "Post-pandemic")))

# Calculate the reductions for Male and Female candidates by period
data_tibble <- data_tibble %>%
  arrange(Date) %>%
  group_by(Period) %>%
  mutate(
    Male_tests_Reduction = c(NA, diff(`Male tests-Conducted`)),
    Female_tests_Reduction = c(NA, diff(`Female tests-Conducted`))
  )

# Remove rows with NA values (first row of each period has NA for reduction)
data_tibble <- data_tibble %>%
  filter(!is.na(Male_tests_Reduction), !is.na(Female_tests_Reduction))

# Fit the linear model for Male tests reduction
model_male <- lm(Male_tests_Reduction ~ Period, data = data_tibble)
summary(model_male)

# Fit the linear model for Female tests reduction
model_female <- lm(Female_tests_Reduction ~ Period, data = data_tibble)
summary(model_female)
```

#-----------------------1.	Hypothesis 1 (Gender-based Focus): Chi-Square Test-----------------------
```{r}
# Create a contingency table for the Chi-square test
# Summarize reductions by gender and period
data_tibble_summary <- data_tibble %>%
  group_by(Period) %>%
  summarise(
    Male_Reduction = sum(Male_tests_Reduction, na.rm = TRUE),
    Female_Reduction = sum(Female_tests_Reduction, na.rm = TRUE)
  )

# Convert reductions into a matrix for the Chi-square test
reduction_matrix <- as.matrix(data_tibble_summary[, c("Male_Reduction", "Female_Reduction")])
rownames(reduction_matrix) <- data_tibble_summary$Period

# Perform the Chi-square test
chi_square_result <- chisq.test(reduction_matrix)

# Output the test result
print(chi_square_result)
```

# Visulize the Hypothesis 01 - Chi-Square Data
```{r}
# Create a dataset for plotting by summarizing Male and Female reductions by period and reshaping the data
ggplot(data_tibble %>%
         group_by(Period) %>%
         summarise(
           Male_Reduction = sum(Male_tests_Reduction, na.rm = TRUE),
           Female_Reduction = sum(Female_tests_Reduction, na.rm = TRUE) 
         ) %>% 
         pivot_longer(cols = c(Male_Reduction, Female_Reduction), names_to = "Gender", values_to = "Reduction") %>%
         mutate(Gender = recode(Gender, "Male_Reduction" = "Male", "Female_Reduction" = "Female")), 
       aes(x = Period, y = Reduction, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +  # Create a dodged bar plot
  scale_fill_manual(values = c("Male" = "skyblue", "Female" = "hotpink")) +  # Correct color mapping
  labs(
    title = "Reduction in Automatic Driving Test Candidates by Gender and Period",
    x = "Period",
    y = "Reduction in Candidates",
    fill = "Gender"
  ) +
  theme_minimal() +  # Apply a minimal theme
  scale_y_continuous(labels = scales::comma)  # Format y-axis with commas for better readability
```

# Hypothesis 01 Discussion - Overall Summary, Linear Modeal VS Chi-Square
The analysis supports Hypothesis 1, showing significant differences in the reduction of driving test candidates between males and females during the COVID-19 pandemic. The Chi-square test revealed a highly significant difference in reductions across gender and periods (p-value < 2.2e-16), indicating that gender-based impacts were not equal. The linear regression models also indicated reductions in both male and female candidates, but due to limited data and perfect model fit, the Chi-square test was more reliable for confirming the significant difference between genders. Thus, the Chi-square test is the more robust method for assessing gender-based differences in reductions.

#-----------------------2.	Hypothesis 2 (Year-based Focus): Linear Model Test-----------------------
```{r}
# Load the data function
load_data <- function() {
  data <- read_excel("./PreparedData.xlsx")
  as_tibble(data)
}

# Reload the data before Hypothesis 2 analysis
data_tibble <- load_data()

# Hypothesis 2: Year-based Focus Analysis
# Recreate Period column to ensure consistency
data_tibble <- data_tibble %>%
  mutate(Period = case_when(
    Date %in% c("2018/19", "2019/20") ~ "Pre-pandemic",
    Date %in% c("2020/21", "2021/22") ~ "Pandemic",
    Date %in% c("2022/23", "2023/24") ~ "Post-pandemic",
    TRUE ~ "Unknown"
  )) %>%
  mutate(Period = factor(Period, levels = c("Pre-pandemic", "Pandemic", "Post-pandemic")))

# Calculate the reduction in 'Total tests-Conducted' for each period
data_tibble <- data_tibble %>%
  arrange(Date) %>%
  group_by(Period, .drop = FALSE) %>%  # Retain unused levels
  mutate(Test_Reduction = c(NA, diff(`Total tests-Conducted`))) %>%
  filter(!is.na(Test_Reduction))  # Remove rows with NA values

# Check levels after filtering
print(levels(data_tibble$Period))
print(table(data_tibble$Period))

# Fit the linear model to examine the reduction in tests by period
model_total_tests <- lm(Test_Reduction ~ Period, data = data_tibble)

# Output the summary of the linear model
summary(model_total_tests)

```


#-----------------------2.	Hypothesis 2 (Year-based Focus): Chi-Square Test-----------------------
```{r}
# Reload the data to ensure a clean slate
data_tibble <- load_data()

# Recreate Period column
data_tibble <- data_tibble %>%
  mutate(Period = case_when(
    Date %in% c("2018/19", "2019/20") ~ "Pre-pandemic",
    Date %in% c("2020/21", "2021/22") ~ "Pandemic",
    Date %in% c("2022/23", "2023/24") ~ "Post-pandemic",
    TRUE ~ "Unknown"
  )) %>%
  mutate(Period = factor(Period, levels = c("Pre-pandemic", "Pandemic", "Post-pandemic")))

# Calculate the reductions in 'Total tests-Conducted' for each period
data_tibble <- data_tibble %>%
  arrange(Date) %>%
  group_by(Period, .drop = FALSE) %>%
  mutate(Test_Reduction = c(NA, diff(`Total tests-Conducted`))) %>%
  filter(!is.na(Test_Reduction))  # Remove rows with NA values

# Summarize reductions by period
data_tibble_summary <- data_tibble %>%
  group_by(Period) %>%
  summarise(Total_Reduction = sum(Test_Reduction, na.rm = TRUE))

# Create a contingency table for Chi-Square Test
reduction_matrix <- as.matrix(data_tibble_summary[, "Total_Reduction"])
rownames(reduction_matrix) <- data_tibble_summary$Period

# Perform the Chi-Square Test
chi_square_result <- chisq.test(reduction_matrix)

# Output the test result
print(chi_square_result)

```

# Visulize the Hypothesis 02 - Chi-Square Data
```{r}
# Visualization of Total Reductions by Period for Hypothesis 2
ggplot(data_tibble_summary, aes(x = Period, y = Total_Reduction, fill = Period)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = c("Pre-pandemic" = "blue", "Pandemic" = "#FFA500", "Post-pandemic" = "#008080")) +
  labs(
    title = "Reduction in Automatic Driving Test Candidates by Period",
    x = "Period",
    y = "Total Reduction in Candidates",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold")
  ) +
  scale_y_continuous(labels = scales::comma)  # Format y-axis with commas for readability
```

# Hypothesis 02 Discussion - Overall Summary, Linear Modeal VS Chi-Square
The analysis supports Hypothesis 2, showing significant differences in the reduction of driving test candidates across the pre-pandemic, pandemic, and post-pandemic periods. The Chi-Square Test revealed a highly significant difference in reductions across these periods (p-value < 2.2e-16), indicating that the impact of the pandemic on driving test candidates was not consistent with the other periods.

The linear regression model attempted to quantify the reductions but was limited by the small sample size, resulting in perfect model fit and indeterminate p-values due to zero degrees of freedom. While the regression model highlighted noticeable differences in reductions across periods, the Chi-Square Test was more robust in confirming these differences.

Thus, the Chi-Square Test provides stronger evidence supporting Hypothesis 2, emphasizing that the reduction in driving test candidates during the pandemic was significantly different from both the pre-pandemic and post-pandemic periods.
